\subsection{Statistical Methods}
\label{sec:monoZ_stats}

%%%%%%%%%%%%%%%%%%%

To reliably derive physical conclusions from the collected experimental data, a robust statistical framework is required. This section provides a concise overview of the specific statistical machinery deployed throughout the DM search program.

Within modern high-energy physics analyses, theoretical postulations are rigorously evaluated against empirical observations via statistical hypothesis testing. This inference methodology is centrally designed to quantify whether the recorded data statistically favor a specific phenomenological model. Conventionally, the baseline background-only paradigm—assuming the complete absence of the targeted new physics signal—is formulated as the null hypothesis ($H_0$). Conversely, the alternative hypothesis ($H_1$) accommodates a signal-plus-background configuration, intrinsically implying the existence of the newly proposed BSM interaction. The ultimate objective of this testing procedure is to quantitatively measure the degree of evidence resisting the null hypothesis.

\subsubsection{The $p$-Value}
\label{sec:p-value}

To quantitatively assess how well the observed experimental outcomes align with a specific theoretical postulate, the $p$-value metric is conventionally employed. Formally, assuming the background-only baseline ($H_0$) holds exactly true, this metric represents the mathematical probability of yielding a test statistic that is equally or more extreme than the actual measurement. Consequently, an observation characterized by a minute $p$-value is deemed highly improbable under the pure-background assumption, thereby furnishing potential evidence in support of the alternative signal hypothesis. Within the particle physics community, such an excess is routinely translated into a Gaussian statistical significance, mapping the derived $p$-value to a corresponding number of standard deviations ($\sigma$).

\subsubsection{Statistical Significance}
\label{sec:significance}

The statistical significance acts as a direct numerical translator for the $p$-value, articulating the degree of incompatibility between the empirical data and $H_0$ using the standard deviation ($\sigma$) scale of a standard normal distribution.

To contextualize, achieving a $2\sigma$ significance is mathematically equivalent to a $p$-value of roughly 0.0455 (4.55\%), implying only marginal tension with the background-only model. Reaching the $3\sigma$ threshold corresponds to a $p$-value of approximately $2.7 \times 10^{-3}$, constituting compelling evidence for an anomaly. Ultimately, a definitive scientific discovery dictates a rigorous $5\sigma$ standard (yielding a $p$-value of roughly $5.7 \times 10^{-7}$). Crossing this threshold essentially guarantees that the observed experimental excess is overwhelmingly unlikely to manifest merely as a random upward statistical fluctuation of the underlying SM background.

\subsubsection{Likelihood Function}
\label{sec:likelihood}

The statistical separation between the background-only and signal-plus-background hypotheses relies heavily on the construction of a robust test statistic. The most ubiquitous choice in collider physics is built upon the \textit{likelihood function}, a mathematical construct that defines the probability of realizing the observed dataset given a postulated set of theory parameters.

Because event counts populating localized phase-space regions inherently obey Poisson statistics, the interaction strength of the targeted physical process is typically controlled by an unconstrained scaling factor, designated as the \textit{signal strength} ($\mu$). This factor scales the nominal expected signal yields and acts as the primary \textit{parameter of interest} (POI) in the fit. A best-fit configuration of $\mu = 0$ perfectly recovers the $H_0$ scenario, whereas a statistically significant positive $\mu$ estimate highlights a measurable signal contribution residing above the background floor.

To accurately model the real-world detector environment, additional degrees of freedom—termed \textit{nuisance parameters} ($\theta$)—are embedded within the likelihood to profile various systematic uncertainties. These parameters are not inherently interesting for the final physics measurement but are vital for correct uncertainty propagation. The full likelihood function, parameterized simultaneously by the POI $\mu$ and the nuisance array $\theta$, is formulated as:

\begin{equation}
\label{eqn:likelihood}
L(\mu,\theta \mid \text{data}) = \prod_{j=1}^{N} \frac{(\mu s_j + b_j)^{n_j}}{n_j!} e^{-(\mu s_j + b_j)},
\end{equation}
where the integer $n_j$ denotes the actual count of observed data events residing in the $j$-th bin. Conversely, $s_j$ and $b_j$ symbolize the theoretically predicted yields for the signal and background processes, respectively, within that identical bin. To extract the optimal signal strength, an unconditional maximization of this likelihood is performed with respect to both $\mu$ and the entire suite of $\theta$ parameters, tightly constrained by the experimental observations.

In sophisticated analyses, evaluating the parameter $\mu$ solely within the signal region is insufficient; dedicated \textit{control regions} are routinely incorporated to anchor the dominant background normalizations. By jointly fitting these regions, the global likelihood expands into:

\begin{equation}
\label{eqn:likelihood2}
L(\mu,\theta \mid \text{data}) = \prod_{j=1}^{N} \frac{(\mu s_j + b_j)^{n_j}}{n_j!} e^{-(\mu s_j + b_j)}
\prod_{k=1}^{M} \frac{m_k^{u_k}}{u_k!} e^{-m_k},
\end{equation}
where the variables $u_k$ and $m_k$ respectively designate the experimentally recorded event counts and the corresponding background expectations within the $k$-th bin of the chosen control regions. Constructing this combined likelihood enables a rigorous, simultaneous profiling procedure that extracts the signal POI while systematically calibrating background estimations directly from data.

\subsubsection{Confidence Intervals}
\label{sec:CL}

When evaluating constraints on specific theoretical parameters or determining upper exclusion bounds on cross-sections, the modified frequentist prescription, universally known as the $\text{CL}_s$ method, is adopted. This rigorous paradigm mitigates the risk of excluding signals to which the analysis possesses no actual sensitivity by evaluating two distinct probabilities: $\text{CL}_{s+b}$ (the confidence level associated with the signal-plus-background hypothesis) and $\text{CL}_b$ (the confidence level of the null hypothesis). The exclusion metric itself is formed via the ratio:
\[
\text{CL}_s = \frac{\text{CL}_{s+b}}{\text{CL}_b}.
\]
Standard operating procedures dictate that a theoretical signal configuration is formally excluded at a $(1-\alpha)$ confidence level whenever the calculated $\text{CL}_s$ falls below the predefined $\alpha$ tolerance. Typically, $\alpha$ is fixed at 0.05, establishing the widely quoted 95\% confidence level limit.

Equation~\eqref{eqn:likelihood2} serves as the foundational core of this statistical machinery. It integrates the pure Poisson counting probabilities across all populated analysis bins with auxiliary penalty constraints that map the intricate network of systematic uncertainties (encompassing theoretical modeling, detector resolutions, and background estimations) onto the $\theta$ nuisance parameters.

For the purpose of estimating the optimal parameter of interest, the \textit{profile likelihood ratio} is overwhelmingly selected as the optimal test statistic. Operating on the signal strength $\mu$, this ratio is mathematically defined as:
\[
\lambda(\mu) =
\frac{L(\mu, \hat{\hat{\theta}}(\mu))}
     {L(\hat{\mu}, \hat{\theta})},
\]
in which $\hat{\hat{\theta}}(\mu)$ represents the conditionally maximized estimates for the nuisance parameters assuming a rigidly fixed $\mu$. In contrast, the denominator utilizes the completely unconditional global maximum-likelihood estimates, denoted as $(\hat{\mu}, \hat{\theta})$, optimized simultaneously over all free parameters.

Relying on this robust likelihood ratio, the primary test statistic is formulated as:
\[
q_\mu = -2 \ln \lambda(\mu)
\]
which serves as the primary metric for quantifying the consistency between the empirical observations and any presumed signal configuration. Thanks to the established asymptotic behaviors of this specific test statistic, one can reliably calculate critical $p$-values and construct confidence intervals analytically, frequently bypassing the need for computationally exhaustive Monte Carlo pseudo-experiments.

Ultimately, the deployment of these sophisticated statistical techniques guarantees a highly coherent and robust framework for hypothesis discrimination, parameter profiling, and theoretical limit derivation, paving the way for sound physical interpretations of the ATLAS dataset.