% !TeX root = ../main.tex

\ustcsetup{
  keywords  = {ATLAS, Z玻色子, 暗物质, 监测漂移管探测器},
  keywords* = {LHC, ATLAS, ZZ, Dark Matter, sMDT},
}

\begin{abstract}

希格斯（Higgs）玻色子被发现后，粒子物理标准模型的理论体系得以进一步完善。当前粒子物理领域的核心研究方向主要包含两大内容：一是精准验证标准模型的理论预言，二是探寻标准模型框架之外的新物理现象。$W$ 玻色子与 $Z$ 玻色子是传递弱核力的基本媒介粒子，自 1983 年被实验发现以来，始终是标准模型的核心组成部分。在大型强子对撞机（LHC）实验中，$ZZ$ 玻色子对产生过程是极为重要且事例占比极高的末态过程，该过程的主要观测衰变道为四轻子衰变道与双轻子加不可见粒子末态衰变道。其中，双轻子加不可见粒子末态具备显著的实验优势：相较于四轻子衰变道，该末态的产生截面更大，能够有效提升实验测量的统计样本量，进而优化测量精度；同时该末态结构简洁，仅包含两个可观测轻子，对应的物理背景清晰、数据分析流程简便，是开展 $ZZ$ 玻色子对产生截面测量与观测的优质衰变道。

在标准模型理论框架下，希格斯玻色子存在极小概率衰变至一对 $Z$ 玻色子并最终进一步衰变为四个中微子，该衰变分支比仅约为 0.001。在暗物质探测及相关研究中，诸多超越标准模型的新物理理论预言了希格斯玻色子的不可见衰变模式。对于希格斯玻色子的不可见衰变（$H\to\text{inv}$），其关联产生过程（$pp\rightarrow ZH \rightarrow \ell^+\ell^- + \text{inv}$）与标准模型 $ZZ\rightarrow \ell^+\ell^- \nu \bar\nu$ 衰变过程类似，实验中仅能观测到单个 $Z$ 玻色子的可见信号，其最终观测末态同样表现为双轻子加不可见粒子组分。基于该衰变道的实验特征，可通过统计测量该末态的事例产额，实现对希格斯玻色子新物理衰变道的探究，并给出其不可见衰变分支比的上限。本文针对 spin-1 新物理模型与 $\text{2HDM}+a$ 双希格斯二重态模型开展研究，结合两类模型的理论预言，给出了更为精确的希格斯玻色子不可见衰变观测上限。

% 本文基于 ATLAS 探测器于 2015--2018 年间采集的 $13 \text{ TeV}$ 质子-质子对撞数据（对应积分亮度达 $140 \text{ fb}^{-1}$），通过系统分析双轻子加不可见粒子末态的实验事例，完成了对 $ZZ$ 玻色子对产生微分截面的精准测量。实验结果表明：全同包容性 $ZZ$ 产生过程的基准截面测量值为 $21.0 \pm 1.0$ fb；在 $Z$ 玻色子质量 $66--116 \text{ GeV}$ 区间内，经全相空间外推得到的总产生截面为 $15.38 \pm 0.81$ pb；同时，伴随双喷注的 $ZZjj$ 产生过程基准截面测量值为 $0.96^{+0.18}_{-0.16}$ fb。相较于 ATLAS 实验第二次运行（Run 2）早期的测量结果，本文包容性基准测量的总不确定度由 $7.0\%$ 降至 $4.6\%$，测量精度显著提升，且所有实验测量结果均与标准模型理论预言高度吻合。此外，本文利用同一批实验数据，探究了该末态下希格斯玻色子发生超越标准模型不可见衰变的可能性。结果表明，现有新物理理论预言的相关信号未在实验中被显著观测到，本文据此重新设定了希格斯玻色子不可见衰变分支比的约束上限。此外，本文利用同一批实验数据，探究了该末态下希格斯玻色子发生超越标准模型不可见衰变的可能性。结果表明，现有新物理理论预言的相关信号未在实验中被显著观测到，本文据此重新设定了希格斯玻色子不可见衰变分支比的约束上限。
% 因字数不满足学校要求，添加更多技术细节来保证最低字数

本文基于 ATLAS 探测器于 2015--2018 年间采集的 $13 \text{ TeV}$ 质子-质子对撞数据（对应积分亮度达 $140 \text{ fb}^{-1}$），通过系统分析双轻子加不可见粒子末态的实验事例，完成了对 $ZZ$ 玻色子对产生微分截面的精准测量。在数据分析与本底压制策略方面，本研究设计了严格的信号区运动学筛选条件（包括设定高缺失横动量 $E_{T}^{\text{miss}}$ 阈值、$E_{T}^{\text{miss}}/H_{T}$ 比例下限以及严格的 b 喷注否决等），以有效抑制截面巨大的 $Z$+jets 和顶夸克（top-quark）本底。为进一步提升测量精度并控制系统误差，研究中针对 $WZ$、$WW$、顶夸克及 $Z$+jets 等主要干扰背景分别构建了专用的独立控制区。此外，基于双轻子系统与缺失横动量之间的方位角差 $\Delta\phi(\ell\ell, \vec{E}_{T}^{\text{miss}})$ 分布，本研究在信号区与多个控制区之间开展了同步轮廓似然拟合（Simultaneous profile-likelihood fit）。这一多区域联合拟合方法实现了对各项主要本底归一化因子的精确约束，极大降低了对蒙特卡洛模拟的依赖。实验结果表明：全同包容性 $ZZ$ 产生过程的基准截面测量值为 $21.0 \pm 1.0 \text{ fb}$；在 $Z$ 玻色子质量 $66\text{--}116 \text{ GeV}$ 区间内，经全相空间外推得到的总产生截面为 $15.38 \pm 0.81 \text{ pb}$；同时，伴随双喷注的 $ZZjj$ 产生过程基准截面测量值为 $0.96^{+0.18}_{-0.16} \text{ fb}$。相较于 ATLAS 实验第二次运行（Run 2）早期的测量结果，本文包容性基准测量的总不确定度由 $7.0\%$ 降至 $4.6\%$，测量精度显著提升，且所有实验测量结果均与标准模型理论预言高度吻合。

此外，本文利用同一批实验数据，探究了该末态下希格斯玻色子发生超越标准模型不可见衰变以及暗物质伴随产生的可能性。在希格斯不可见衰变的搜寻中，本研究引入了提升多变量决策树（BDT）机器学习算法，将多个运动学变量组合以最大化信号与本底的分离度；而在简化暗物质模型与 $\text{2HDM}+a$ 模型的搜寻中，则采用 $ZZ$ 系统的横质量（$m_{T}$）分布作为核心变量进行拟合。结果表明，实验数据中并未显著观测到现有新物理理论预言的异常信号。基于这个结果，在假设希格斯玻色子产生截面与标准模型预期一致的前提下，本文在 $95\%$ 置信度（CL）下，将希格斯玻色子不可见衰变分支比 $\mathcal{B}(H \to \text{inv})$ 的观测（预期）上限严格限制在了 $18.7\%$（$19\%$）。同时，针对简化暗物质模型和包含额外赝标量媒介子的双希格斯二重态模型（$\text{2HDM}+a$），本研究都给出了更为精确的边界。最后，本文将对撞机框架下的暗物质搜索结果转化为弱相互作用大质量粒子（WIMP）与核子及质子的散射截面限制，并与最新的探测实验结果进行了系统比较。研究表明，对撞机探测与直接探测途径形成了显著的物理互补，本测量在低暗物质质量区间展现出了相当的探测优势。

当前，大型强子对撞机正逐步升级为高亮度大型强子对撞机（HL-LHC）。更高的对撞亮度将为新物理信号的挖掘与新物理现象的发现提供更多机遇，但同时也对探测器的抗辐照性能、事例计数率耐受能力等核心指标提出了更为严苛的要求，探测器硬件系统的迭代升级不可或缺。本文最后阐述了作者在 ATLAS 探测器升级项目中的硬件研究工作，主要包括小型监测漂移管（sMDT）探测器的组装、调试与性能测试。这些新型探测器将应用于 ATLAS 探测器的整体升级改造中，以保障高亮度对撞环境下的粒子探测性能，为大型强子对撞机下一步的高精度物理运行奠定坚实的硬件基础。

\end{abstract}

%中文摘要是论文内容的总结概括，应简要说明论文的研究目的、基本研究内容、研究方法或过程、结果和结论，突出论文的创新之处。摘要应具有独立性和自明性，即不用阅读全文，就能获得论文必要的信息。摘要中不宜使用公式、图表，不引用文献。博士论文摘要一般控制在800-1500字
%中文关键词是为了文献标引工作从论文中选取出来用以表示全文主题内容信息的单词和术语，一般 3～8 个词，要求能够准确概括论文的核心内容。

\begin{abstract*}

This thesis utilizes data collected by the ATLAS detector from proton-proton collisions at a center-of-mass energy of $\sqrt{s} = 13\text{ TeV}$ between 2015 and 2018, corresponding to an integrated luminosity of $140\text{ fb}^{-1}$. The analyses focus on the di-lepton plus invisible particles final state, systematically investigating both the Standard Model (SM) $ZZ$ boson pair production and the potential beyond the Standard Model (BSM) invisible decays of the Higgs boson and associated dark matter production.

In terms of data analysis and background suppression strategies, strict kinematic selection criteria in the signal region are designed, including high missing transverse momentum $E_{T}^{\text{miss}}$ thresholds, an $E_{T}^{\text{miss}}/H_{T}$ ratio lower limit, and a stringent b-jet veto. These selections effectively suppress the massive backgrounds from $Z$+jets and top-quark processes. To further improve measurement precision and control systematic uncertainties, dedicated independent control regions are constructed for the main background sources, including $WZ$, $WW$, top-quark, and $Z$+jets processes. Furthermore, based on the azimuthal angle difference between the di-lepton system and the missing transverse momentum vector, $\Delta\phi(\ell\ell, \vec{E}_{T}^{\text{miss}})$, a simultaneous profile-likelihood fit is performed across the signal region and multiple control regions. This multi-region joint fitting method successfully constrains the normalization factors of the major backgrounds, significantly reducing the dependence on Monte Carlo simulations.

Based on this robust methodology, a precise measurement of the $ZZ$ boson pair production differential cross-sections has been achieved. The experimental results show that the measured fiducial cross-section for the total inclusive $ZZ$ production process is $21.0 \pm 1.0\text{ fb}$. Within the $Z$ boson mass range of $66\text{--}116\text{ GeV}$, the total production cross-section extrapolated to the full phase space is $15.38 \pm 0.81\text{ pb}$. Meanwhile, the measured fiducial cross-section for the $ZZ$ production in association with two jets ($ZZjj$) is $0.96^{+0.18}_{-0.16}\text{ fb}$. Compared to the early measurements from the ATLAS Run 2, the total uncertainty of the inclusive fiducial measurement is reduced from $7.0\%$ to $4.6\%$, representing a significant precision improvement. All experimental measurements are in excellent agreement with the SM predictions, including advanced calculations incorporating next-to-next-to-leading order (NNLO) QCD and NLO electroweak corrections. Additionally, by exploiting the $Z$ boson transverse momentum ($p_T^Z$) and the $ZZ$ system transverse mass ($m_T^{ZZ}$) distributions, stringent constraints are placed on anomalous neutral triple gauge couplings (aTGCs) and anomalous quartic gauge couplings (aQGCs) within both the vertex function approach and the Standard Model Effective Field Theory (SMEFT) frameworks.

Furthermore, using the same dataset, this thesis investigates the possibility of the Higgs boson decaying into BSM invisible modes and the associated production of dark matter. In the search for Higgs invisible decays, a Boosted Decision Tree (BDT) machine learning algorithm is introduced to combine multiple kinematic variables to maximize the separation between signal and background. In the searches for simplified dark matter models and the Two-Higgs-Doublet Model with an additional pseudoscalar mediator ($\text{2HDM}+a$), the transverse mass ($m_{T}$) distribution of the $ZZ$ system is adopted as the core variable for fitting. The results indicate that no significant excesses over the SM expectations are observed in the data. Based on this outcome, assuming the Higgs boson production cross-section is consistent with the SM expectation, the observed (expected) upper limit on the Higgs boson invisible decay branching ratio $\mathcal{B}(H \to \text{inv})$ is strictly constrained to $18.7\%$ ($19\%$) at $95\%$ confidence level (CL). Concurrently, more precise boundary constraints are set for the probed dark matter models.

Finally, the dark matter search results from the collider framework are translated into limits on the scattering cross-sections of Weakly Interacting Massive Particles (WIMPs) with nucleons and protons. These limits are systematically compared with the latest direct detection experiments. The study demonstrates that collider searches and direct detection approaches exhibit significant physical complementarity, and this measurement reveals a substantial detection advantage in the low dark matter mass region.

In addition to the detailed physics analyses presented, this thesis also highlights a significant hardware project related to the ATLAS muon detector upgrade, a critical task in preparation for the high-luminosity phase of the Large Hadron Collider (HL-LHC). 

The focus of this work lies in the production and testing of the small-diameter Monitored Drift Tube (sMDT) chambers. The author's contributions span various aspects of this undertaking, including the development and implementation of infrastructure for detector construction, the meticulous assembly of precision tubes and chambers, and the rigorous execution of quality-control tests. These efforts are integral to ensuring the ATLAS experiment's readiness to capitalize on the enhanced physics potential afforded by the HL-LHC era.



%
%  Following the discovery of the Higgs boson, the Standard Model(SM) of particle physics has become increasingly robust. However, verifying the SM and exploring the beyond Standard Model(BSM) theories remains the primary focus in particle physics research. W and Z bosons are fundamental particles responsible for mediating the weak force, and they have been pivotal components of the Standard Model since their discovery in 1983. Particularly, the production of ZZ boson pairs is a significant and prevalent channel in the Large Hadron Collider (LHC). Its primary visible decay channels, including four leptons or two leptons accompanied by 2 invisible neutrinos, provide valuable insights for the verification of SM. Specifically, the di-lepton plus invisible final state has a significantly higher cross-section compared to the four-lepton channel, enhancing statistical significance and measurement precision. Moreover, its simplicity in the final state and physics process could facilitate the analysis.

% In the Standard Model theory, the Higgs boson has a small branching ratio (~0.001) to decay into two Z bosons, then to four neutrinos. However, some BSM(beyond the Standard Model) theories predict other invisible decays of the Higgs boson, which can be used for dark matter search. Therefore the ZH production in such scenarios, can only observe one visible Z boson(decays to visible state $ll$) in the final state, accompanied by an invisible part. By measuring the cross-section of events in this final state, we can probe the existence of Higgs boson decays with BSM predictions and set limitations on their branching ratios. 

%This thesis utilizes data collected by the ATLAS detector at $\sqrt{s}=13 TeV$, corresponding to an integrated luminosity of $140 fb^{-1}$ over the period 2015-2018. Through analysis and measurement of the $ll+E^{miss}_{T}$ final state, the differential cross-section for ZZ boson pair production is measured with the latest data. The obtained results show higher precision compared to previous measurements with early run-2 data corresponding to an integrated luminosity of $36.1 fb^{-1}$. The second analysis investigates the invisible decays of Higgs boson beyond the Standard Model predictions,  no significant excesses were observed above the SM expectation, and limits are set on the signal models.

%In addition to the detailed physics analyses presented, this thesis also highlights a significant hardware project related to the ATLAS muon detector upgrade, a critical task in preparation for the high-luminosity phase of the Large Hadron Collider (HL-LHC) expected in 2030. 
%The focus of this work lies in the production and testing of the small-diameter Monitored Drift Tube (sMDT) chambers. The author's contributions span various aspects of this undertaking, including the development and implementation of infrastructure for detector construction, the meticulous assembly of precision tubes and chambers, and the rigorous execution of quality-control tests. These efforts are integral to ensuring the ATLAS experiment's readiness to capitalize on the enhanced physics potential afforded by the HL-LHC era.
%


 
\end{abstract*}


