% !TeX root = ../main.tex

\begin{publications}

\section*{已发表论文 / Published Papers}

\begin{enumerate}

\item \textbf{[Chuanshun Wei]}, [A.Chen, D.Amidei, N.Anderson, E.Carpenter, L.Cooperrider, T.Dai, E.Diehl, C.Ferretti, J.Li, P.Lugato, J.Minnella, S.Moskaitis, K.Nelson, V.Pillsbury, E.Salzer, L.Simpson, Z.C.Wang, Z.R.Wang, C.Weaverdyck, Z.Yang, B.Zhou], et al. ``Construction and testing of sMDT tubes at the University of Michigan for the ATLAS Muon Spectrometer upgrade.'' \textit{Journal of Instrumentation (JINST)}, 17, P10010 (2022). DOI: 10.1088/1748-0221/17/10/P10010.
\\ \textbf{Personal Contribution:} [Primary author. Main member in construction and tests]

\item \textbf{[Chuanshun Wei]}, [D. Amidei, N. Anderson, A. Chen, E. Carpenter, L. Cooperrider, T. Dai, E. Diehl, C. Ferretti, Y. Guo, J. Li, X. Meng, K. Nelson, V. Pillsbury, E. Salzer, T. Schwarz, L. Simpson, Z. Wang, C. Weaverdyck, Z. Yang, M. Yuan, B. Zhou, J. Zhu], et al. ``Construction of precision sMDT detector for the ATLAS Muon Spectrometer upgrade.'' \textit{Journal of Instrumentation (JINST)}, 18, P01041 (2023). DOI: 10.1088/1748-0221/18/01/P01041.
\\ \textbf{Personal Contribution:} [Primary author. Main member in construction and tests]

\item ATLAS Collaboration. ``Search for associated production of a $Z$ boson with an invisibly decaying Higgs boson or dark matter candidates at $\sqrt{s}=13$~TeV with the ATLAS detector.'' \textit{Physics Letters B}, 829, 137066 (2022). DOI: 10.1016/j.physletb.2022.137066.
\\ \textbf{Personal Contribution:} [Primary author. ZZ systematics, code developer for ntuples]

\item ATLAS Collaboration. ``Combination and summary of ATLAS dark matter searches interpreted in a 2HDM with a pseudo-scalar mediator using $139~\text{fb}^{-1}$ of $\sqrt{s}=13$~TeV $pp$ collision data.'' \textit{Science Bulletin}, 69(19), 3005-3035 (2024). DOI: 10.1016/j.scib.2024.06.003.
\\ \textbf{Personal Contribution:} [Primary author. Orthogonality check, several summary plot]

\end{enumerate}

\section*{待发表论文 / Papers to be Published}

\begin{enumerate}
\item ATLAS Collaboration. ``Measurements of $ZZ \rightarrow \ell\ell\nu\nu$ and $ZZjj \rightarrow \ell\ell\nu\nu jj$ productions in $pp$ collisions at $\sqrt{s}=13$~TeV with the ATLAS detector.'' arXiv:2511.15569 [hep-ex] (2025). 
\\ \textbf{Personal Contribution:} [Primary author. CRs validation, DNN approach, QCD/PDF, Inclusive Fit, ALternative unfolding]
\end{enumerate}

\section*{研究报告 / Research Reports}
\begin{enumerate}
\item 2021 APS April Meeting, sMDT Module-0 Chamber Construction and Test for HL-LHC Upgrade of the ATLAS Muon Spectrometer. 
\item 2021 DPF Meeting, Search for invisible decays of the Higgs boson and for dark matter with the ATLAS detector. 
\item 2023 APS April Meeting, Cross section measurement for the production of ZZ → llvv with the ATLAS detector. 
\end{enumerate}

\end{publications}
