% !TeX root = ../main.tex

\chapter{\texorpdfstring{Measurement of ZZ Production with $ll\nu\nu$ Final State}{Measurement of ZZ Production with llnunu Final State}}


\section{Introduction}
\label{sec:ZZ_intro}

The study of diboson production at the Large Hadron Collider (LHC) is a cornerstone of the experimental program to test the Standard Model (SM) of particle physics. Among these processes, the production of $Z$ boson pairs ($ZZ$) is of particular importance. It provides a direct probe of the electroweak gauge structure and the non-Abelian self-interactions of the gauge bosons. Precise measurements of the $ZZ$ production cross-section constitute a stringent test of the SM predictions at the highest available energies and serve as a sensitive probe for new physics phenomena that may manifest as anomalous Triple Gauge Couplings (aTGCs).

This chapter presents a measurement of the $ZZ$ production cross-section in proton-proton collisions at a center-of-mass energy of $\sqrt{s} = 13$ TeV, using the full Run 2 dataset collected by the ATLAS detector, corresponding to an integrated luminosity of $140~\text{fb}^{-1}$. The analysis focuses on the final state where one $Z$ boson decays into a pair of charged leptons ($\ell^+\ell^-$, where $\ell = e, \mu$) and the other decays into a pair of neutrinos ($\nu\bar{\nu}$).

\subsection{Motivation for the $\ell\ell\nu\nu$ Channel}

While the "golden channel" $ZZ \to 4\ell$ offers a fully reconstructed final state with very high signal purity, it suffers from a small branching ratio. The branching fraction of the $Z$ boson to neutrinos ($\mathcal{B}(Z \to \nu\bar{\nu}) \approx 20\%$) is approximately six times larger than that to charged leptons ($\mathcal{B}(Z \to \ell^+\ell^-) \approx 3.3\%$). Consequently, the $ZZ \to \ell\ell\nu\nu$ channel benefits from significantly higher statistics.

This statistical advantage is particularly crucial in the high transverse momentum ($p_{\mathrm{T}}$) regime. Since deviations from the SM due to effective field theory (EFT) operators or new heavy resonances are expected to be enhanced at high energies, the $\ell\ell\nu\nu$ channel provides superior sensitivity for constraining aTGCs compared to the $4\ell$ channel. Furthermore, understanding the $ZZ \to \ell\ell\nu\nu$ process is essential as it constitutes a major irreducible background for searches for Dark Matter (Mono-$Z$) and high-mass Higgs boson decays.

\subsection{Analysis Strategy and Outline}

The measurement is performed in a fiducial phase space defined to match the experimental acceptance. The analysis relies on identifying events with a dilepton pair consistent with a $Z$ boson mass and significant missing transverse momentum ($E_{\mathrm{T}}^{\text{miss}}$) arising from the neutrinos.

The structure of this chapter is as follows:
\begin{itemize}
    \item Section~\ref{sec:llvv_samples} describes the data and Monte Carlo (MC) simulation samples used to model the signal and background processes.
    \item Section~\ref{sec:llvv_selection} details the reconstruction of physics objects and the selection criteria defining the Signal Regions (SR).
    \item Section~\ref{sec:llvv_background} discusses the estimation of major backgrounds, particularly the dominant $WZ$ and non-resonant backgrounds, using data-driven control regions.
    \item Section~\ref{sec:llvv_systematics} outlines the sources of experimental and theoretical systematic uncertainties.
    \item Section~\ref{sec:llvv_unfolding} presents the unfolding procedure used to correct the detector-level distributions to the particle level.
    \item Section~\ref{sec:llvv_dnn} describes the optimization strategies, including the use of Deep Neural Networks (DNN), to enhance the separation between signal and background.
    \item Section~\ref{sec:llvv_results} reports the measured integrated and differential cross-sections for the inclusive $ZZ$ production and the constraints on aTGCs.
    \item Finally, Section~\ref{sec:llvvjj} extends the analysis to the electroweak production of $ZZ$ in association with two jets ($ZZjj$), targeting the Vector Boson Scattering (VBS) topology and Anomalous Quartic Gauge Couplings (aQGCs).
\end{itemize}


\clearpage
%\section{Data and Monte Carlo Samples}
\input{chapters/4_1_samples.tex}


\clearpage
%\section{Event Selection}
\input{chapters/4_2_events.tex}


\clearpage
%\section{Background Estimation}
\input{chapters/4_3_background.tex}


\clearpage
%\section{Systematic Uncertainty}
\input{chapters/4_4_syst.tex}


\clearpage
%\section{Unfolding}
\input{chapters/4_5_unfolding.tex}


\clearpage
%\section{DNN Optimization}
\input{chapters/4_7_dnn.tex}


\clearpage
%\section{Results}
\input{chapters/4_6_results.tex}


\clearpage
%\section{llvvjj}
\input{chapters/4_8_llvvjj.tex}