% !TeX root = ../main.tex

\chapter{\texorpdfstring{Search for Dark Matter Produced in Association with a $Z$ Boson in the $\ell\ell + E^{\text{miss}}_{\text{T}}$ Final State}{Search for Dark Matter in the ll + MET Final State}}
\label{chap:monoZ}

\section{Introduction}
\label{sec:monoZ_intro}

The existence of Dark Matter (DM) remains one of the strongest evidences for physics Beyond the Standard Model (BSM). While astrophysical observations—ranging from galaxy rotation curves to the Cosmic Microwave Background anisotropy—provide compelling gravitational evidence for DM, its particle nature remains elusive~\cite{Bertone:2004pz}. The Large Hadron Collider (LHC) offers a unique laboratory to produce DM particles in high-energy proton-proton collisions. If DM particles interact weakly with the Standard Model (SM) sector, they may be produced in the ATLAS detector. Being electrically neutral and stable, they would escape detection, manifesting only as a significant imbalance in transverse momentum, denoted as missing transverse momentum ($E_{\mathrm{T}}^{\text{miss}}$)~\cite{Abercrombie:2015wmb}.

To identify such events, the DM particles must be produced in association with a visible SM particle ("X") to recoil against, resulting in a distinct "Mono-X" signature. This chapter presents a search for Dark Matter produced in association with a $Z$ boson, where the $Z$ boson decays into a pair of charged leptons ($\ell^+\ell^-$, with $\ell = e, \mu$). This is referred to as the \textit{Mono-Z} channel~\cite{Aaboud:2017bja}.

This analysis builds directly upon the measurement of the SM $ZZ \to \ell\ell\nu\nu$ process presented in Chapter~\ref{chap:llvv_analysis}. Both analyses share the same final state topology: an on-shell $Z$ boson and large $E_{\mathrm{T}}^{\text{miss}}$. However, they probe distinct kinematic regimes:

\begin{itemize}
    \item \textbf{Precision Measurement:} In the previous chapter, $ZZ \to \ell\ell\nu\nu$ was the signal. The focus was on the inclusive production cross-section and unfolding kinematic spectra.
    \item \textbf{BSM Search:} In this chapter, $ZZ \to \ell\ell\nu\nu$ represents the dominant, irreducible \textbf{background}. The search strategy is optimized for the high-$E_{\mathrm{T}}^{\text{miss}}$ tails where BSM signals are expected to deviate from the SM prediction.
\end{itemize}

The results are interpreted within the context of two benchmark BSM scenarios recommended by the LHC Dark Matter Working Group:
\begin{enumerate}
    \item \textbf{Simplified Models:} Models involving vector or axial-vector mediators (spin-1) that couple the SM quarks to Dirac fermion DM candidates~\cite{Abercrombie:2015wmb}.
    \item \textbf{2HDM+$a$ Models:} A more complex extension involving Two-Higgs-Doublet Models with an additional pseudoscalar mediator ($a$), which provides a rich phenomenology including mono-$Z$ signatures~\cite{Bauer:2017ota}.
\end{enumerate}

The structure of this chapter is as follows. Section~\ref{sec:monoZ_samples} details the simulated samples used to model DM signals and SM backgrounds. Section~\ref{sec:monoZ_selection} describes the event selection criteria, optimized to suppress the $Z+\text{jets}$ and diboson backgrounds while retaining sensitivity to DM signals using the transverse mass of the $ZZ$ system ($m_T^{ZZ}$) as the primary discriminant. The background estimation strategy, utilizing specific Control Regions (CRs) to constrain dominant processes, is discussed in Section~\ref{sec:monoZ_bkg}. Section~\ref{sec:monoZ_syst} outlines the experimental and theoretical systematic uncertainties. Finally, the statistical interpretation and exclusion limits on mediator masses are presented in Section~\ref{sec:monoZ_results}.

\clearpage
% \section{Data and Simulated Samples}
% \label{sec:monoZ_samples}
\input{chapters/5_1_samples.tex}

\clearpage
% \section{Event Selection}
% \label{sec:monoZ_selection}
\input{chapters/5_2_events.tex}

\clearpage
% \section{Background Estimation}
% \label{sec:monoZ_bkg}
\input{chapters/5_3_background.tex}

\clearpage
% \section{Systematic Uncertainties}
% \label{sec:monoZ_syst}
\input{chapters/5_4_syst.tex}

\clearpage
% \section{Statistical Interpretation}
% \label{sec:monoZ_stats}
\input{chapters/5_5_stats.tex}

\clearpage
% \section{Results}
% \label{sec:monoZ_results}
\input{chapters/5_6_results.tex}

