% !TeX root = ../main.tex

\begin{acknowledgements}

First and foremost, my deepest and most profound gratitude goes to my advisors, Professor Zhengguo Zhao and Professor Bing Zhou. Their mentorship, patience, and profound scientific insights have been the bedrock of my academic growth. 

During my time in the United States, Prof. Bing Zhou is an extraordinary guide. I am deeply thankful for her hands-on mentorship and tireless support in both the complex physics analysis and the hardware work for the sMDT chambers construction. Navigating a PhD in a foreign country is challenging in its own right, but doing so during the unprecedented global disruption of the COVID-19 pandemic added immense pressure. Prof. Zhou's steadfast encouragement, patient communication, and clear guidance kept my research on track and provided a beacon of stability during a deeply uncertain time. 

I am equally indebted to Prof. Zhengguo Zhao. A PhD is rarely a linear path, and there were times when I found myself lost, struggling to find my direction and target amidst the immense complexities of high-energy physics research. During this swamp of uncertainty, Prof. Zhao invested an incredible amount of time, effort, and care to help me navigate my way out. His wisdom, sharp scientific intuition, and compassionate guidance not only restored my confidence but also redefined my academic trajectory. I will forever cherish his patience and the profound trust he placed in me to find my footing again. 

I would also like to extend my sincere appreciation to my senior colleagues, Dr. Jing Li and Dr./Prof. Zirui Wang. The two physics analyses presented in this thesis---the precision measurement of the $ZZ$ production cross-section and the search for Dark Matter in the mono-$Z$ channel---would not have reached the current state without their invaluable assistance. They generously shared their experience, spent countless hours troubleshooting the project code, and patiently answered my questions. Their great helped a lot  my transition from a novice student to an independent researcher, and I am grateful for the collaborative spirit. 

I also wish to thank the wider ATLAS collaboration, the Department of Physics at University of Michigan(UMich) and the Department of Particle Physics at the University of Science and Technology of China (USTC). The collaborative environment, the brilliant minds I have interacted with, and the resources provided by the universities have profoundly improved my understanding of physics. 

Finally, I owe a debt of gratitude to my family that words can scarcely capture. To my dear parents and sister, thank you for your endless sacrifices, your unconditional love, and for always believing in my potential, even when my research took me thousands of miles away from home. You have all been my strongest pillar of support, and I share this milestone, and the promising future it brings, with you. 

As I reach the goal of my seven-year doctoral journey, this dissertation is never an individual achievement, but a collective milestone made possible by the unwavering support, guidance, and encouragement of many exceptional individuals. Sincerely, thank you all for the supports! 

\end{acknowledgements}
